---
title: "Model review"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = normalizePath('..'))
knitr::opts_chunk$set(root.dir = normalizePath('..'))
```

```{r load, include=FALSE}
rm(list = ls())
library(cowplot)
source("functions/trimSims.R")
source("functions/SimulationAndPlots.R")
source("multi-neighbor/sim-model-5-data-prep.R")
rm(N_it, N_pop, quarter, case_summary_combined, combined, S_reps, week_date)
files <- dir("data/Rdata/", pattern = "sim._full_data.Rdata", full.names = TRUE)
loadModels <- function(){
  files <- dir("data/Rdata/", pattern = "sim._full_data.Rdata", full.names = TRUE)
  for (i in seq_along(files)){
    
    load(file = files[i])
  }
  rm(files, i)
  browser()
  out <- list(sim)
  return(out)
}

loadModels()

load(file = files)

load(file =  "data/Rdata/sim1_full_data.Rdata")
load(file =  "data/Rdata/sim2_full_data.Rdata")
load(file =  "data/Rdata/sim3_full_data.Rdata")
load(file =  "data/Rdata/sim5_full_data.Rdata")

```
## Full simulation -- all models

* Simulating models 1 to 5 from t = 0.

```{r full, include=FALSE}
sim1_trim <- trimSims(sim1_full_data, 0)
sim2_trim <- trimSims(sim2_full_data, 0)
sim3_trim <- trimSims(sim3_full_data, 0)
sim5_trim <- trimSims(sim5_full_data, 0)

sim1_summary <- SimCI(sim1_trim$sim_filtered)
sim2_summary <- SimCI(sim2_trim$sim_filtered)
sim3_summary <- SimCI(sim3_trim$sim_filtered)
sim5_summary <- SimCI(sim5_trim$sim_filtered)


sim1_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "black", alpha_sim = 0.05,
                     ci = sim1_summary,
                     ribbon = TRUE)
sim2_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "darkgreen", alpha_sim = 0.05,
                     ci = sim2_summary,
                     ribbon = TRUE)
sim3_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "blue", alpha_sim = 0.05,
                     ci = sim3_summary,
                     ribbon = TRUE)
sim5_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "red", alpha_sim = 0.05,
                     ci = sim5_summary,
                     ribbon = TRUE)


```

```{r plot1, echo=FALSE, out.width="1200px", fig.height=10}
plot_grid(sim1_plot, sim2_plot, sim3_plot, sim5_plot,
          labels = c("1", "2", "3", "5"), ncol = 2, nrow = 2)
```


## Constrained models

* Constraining output to only those simulations that achieve at least 10 cummulative infections by day 21 in the St. Annae Vester quarter shows  Model 5 (**red**) and model 2 (**green**) to perform well.  
```{r trimmed, include=FALSE}
sim1_trim <- trimSims(sim1_full_data, 10,
                      quarter_val = "St. Annae Vester", day_val = 21)
sim2_trim <- trimSims(sim2_full_data, 10,
                      quarter_val = "St. Annae Vester", day_val = 21)
sim3_trim <- trimSims(sim3_full_data, 10,
                      quarter_val = "St. Annae Vester", day_val = 21)
sim5_trim <- trimSims(sim5_full_data, 10,
                      quarter_val = "St. Annae Vester", day_val = 21)

sim1_summary <- SimCI(sim1_trim$sim_filtered)
sim2_summary <- SimCI(sim2_trim$sim_filtered)
sim3_summary <- SimCI(sim3_trim$sim_filtered)
sim5_summary <- SimCI(sim5_trim$sim_filtered)


sim1_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "black", alpha_sim = 0.05,
                     ci = sim1_summary,
                     ribbon = TRUE)
sim2_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "darkgreen", alpha_sim = 0.05,
                     ci = sim2_summary,
                     ribbon = TRUE)
sim3_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "blue", alpha_sim = 0.05,
                     ci = sim3_summary,
                     ribbon = TRUE)
sim5_plot <- SimPlot(observed_data = I_reps_plot,
                     rib_col = "red", alpha_sim = 0.05,
                     ci = sim5_summary,
                     ribbon = TRUE)

```

```{r plot2, echo=FALSE, out.width="1200px", fig.height=10}
plot_grid(sim1_plot, sim2_plot, sim3_plot, sim5_plot,
          labels = c("1", "2", "3", "5"), ncol = 2, nrow = 2)


```

***

# Model 5 examination

* Simulate a few from the join posterior
```{r mod5}
sim5_param <- SimFromZero(10,
                          I_reps = I_reps, N_it = N_it,
                          betas_95hpd = mcmc_out$betas_95hpd,
                          phi_95hpd = mcmc_out$phi_95hpd,
                          gamma_95hpd = mcmc_out$gamma_95hpd, seed = 14)

sim_plot <- sim5_param %>%
  SimDataToPlot() %>%
  SimPlotReps(., I_reps_plot, alpha_sim = 1)

```
```{r plo3}
sim_plot
```

## Single parameter combination

* Simulate using the single parameter combination that provided the best result above
* Timing of the outbreak is variable
```{r simparams, fig.width=}
# SIMULATE FROM PARAM POINT VALUES -------------------------------------------------------
param <- sim5_param$store_param
best_param <- lapply(param, function(x) x[[6]])

sim5_timing <- SimFromZeroPointValue(25,
                                     I_reps = I_reps, N_it = N_it,
                                     betas_95hpd = best_param$beta,
                                     phi_95hpd = best_param$phi,
                                     gamma_95hpd = best_param$gamma,
                                     seed = 14)
sim_plot <- sim5_timing %>%
  SimDataToPlot() %>%
  SimPlotReps(., I_reps_plot, alpha_sim = 1) + ggtitle ("model 5: Param point values")
```
```{r plo4}
sim_plot
```

