---
title: "Cholera transmission"
author: "Matthew"
date: "February 9, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

#1 Model
## 1.1 Model structure  
We created five models, each with differining levels of allowed complexity. Each model was based upon the following construction:
$$^{new}\textrm{I}_{i,t+1}\sim Poisson(\lambda_{i,t})$$
$$\lambda_{i,t} = \frac{S_{i,t}}{N_{i}} (\beta_{i}I_{i,t} + \sum_{j\neq i}I_{j,t}\alpha_{j,i} )$$

where:  
$^{new}\textrm{I}_{i,t}$ = the number of new infectious cases in each quarter $i$ at each timepoint $t$  
$I_{i,t}$ = the total number of infectious cases in each quarter $i$ at each timepoint $t$
$S_{i,t}$ = the number of susceptible people in each quarter $i$ at each timepoint $t$  
$N_{i}$ = the total poulation of quarter $i$  
$\beta_{i}$ = the force of internal infection in quarter $i$  


The total number of cases $I_{i,t}$ was updated via:
$$I_{i,t+1} = I_{i,t} + \frac{^{new}\textrm{I}_{i,t}}{\phi} - R_{i,t}$$  
where:  
$\phi$ = the fraction of infectious cases that are reported  
$R_{i,t}$ = the number of people who recovered or died from infection.

The number of recovered individuals $R_{i,t}$ was updated via:  
$$R_{i,t} = \frac{1}{\gamma} I_{i,t}$$
where:  
$\gamma$ = the average duration of infectiousness (fixed at 5 days).  

The number of suceptibles $S_{i,t}$ was updated via:
$$S_{i,t+1} = S_{i,t} - \frac{^{new}\textrm{I}_{i,t}}{\phi}$$

## 1.2 Model priors  


```{r pressure, echo=FALSE}
plot(pressure)
```
