---
title: "The importance of thinking beyond the water-supply in cholera epidemics: a historical urban case-study"
output: pdf_document
author: "Matthew D. Phelps, Andrew S. Azman, Joseph A. Lewnard, Marina Antillón, Lone Simonsen, Viggo Andreasen,  Peter K.M. Jensen, Virginia E. Pitzer."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')
pdf.options(encoding='ISOLatin2')
knitr::opts_knit$set(root.dir = normalizePath('..'))
knitr::opts_chunk$set(root.dir = normalizePath('..'))
library(tidyverse)
library(pander)
```

#S1 Supplemental Text 
## 1.1 Model structure  
We constructed a series of nested models where the force of infection acting upon neighborhood $i$ was the sum of an internal force of infection, $\beta_i$ and external force of infection from neigborhood $j$ upon neighborhood $i$, $\alpha_{j,i}$. 

From simple to complex we allowed (1) a single $\beta$ and single $\alpha$ for
all neighborhoods, such that $\beta_i=\beta$ and $\alpha_{j,i}=\alpha$. (2) An individual $\beta_{i}$ for each neighborhood and a single $\alpha$ for all neighborhoods, such that $\alpha_{j,i}=\alpha$. (3) An individual $\beta_{i}$ for each neighborhood and a single asymmetric $\alpha_{i,j}$ for each neighborhood
pair, such that $\alpha_{j,i} \ne \alpha_{i,j}$.


Two additional models were constructed that were not reported in the main text. In model 2.1 allowed an individual $\beta_{i}$ and $\alpha_{i}$ for each neighborhood, such that $\alpha_{j,i} = \alpha_i$. In model 2.2 we allowed an individual $\beta_{i}$ for each neighborhood and symmetric $\alpha_{j,i}$ for
each neighborhood pair, such that $\alpha_{j,i} = \alpha_{i,j}$. These models were not
supported by the model selection process.

Each model was based upon the following construction:
$$^{new}\textrm{I}_{i,t+1}\sim Poisson \left ( \frac{S_{i,t}\phi}{N_{i}} (\beta_{i}I_{i,t} + \sum_{j\neq i}\alpha_{j,i}I_{j,t} ) \right )$$


where:  
$^{new}\textrm{I}_{i,t}$ = the number of reported new infectious cases in each neighborhood $i$ at time $t$  
$I_{i,t}$ = the total number of infectious cases in each neighborhood $i$ at time $t$  
$S_{i,t}$ = the number of susceptible people in each neighborhood $i$ at time $t$  
$N_{i}$ = the total population of neighborhood $i$  
$\beta_{i}$ = the force of internal infection in neighborhood $i$  
$\alpha_{j,i}$ = the force of infection from neighborhood $j$ to neighborhood $i$.  
$\phi$ = the fraction of cases that are reported  

The total number of cases $I_{i,t}$ was updated via:
$$I_{i,t+1} = I_{i,t} + \frac{^{new}\textrm{I}_{i,t}}{\phi} - R_{i,t}$$  
where $R_{i,t}$ = the number of people who recovered or died from infection.

The number of recovered individuals $R_{i,t}$ was updated via:  
$$R_{i,t+1} = \gamma I_{i,t}$$
where $\frac{1}{\gamma}$ = the duration of infectiousness.  

The number of susceptible $S_{i,t}$ was updated via:
$$S_{i,t+1} = S_{i,t} - \frac{^{new}\textrm{I}_{i,t}}{\phi}$$  

The full system of model equations is thus:
$$S_{i,t+1} = S_{i,t} - \frac{^{new}\textrm{I}_{i,t}}{\phi}$$
$$I_{i,t+1} = I_{i,t} + \frac{^{new}\textrm{I}_{i,t}}{\phi} - R_{i,t}$$
$$R_{i,t+1} = \gamma I_{i,t}$$


## 1.2 Hydraulic connectivity and geographic proximity  
To assess the effect of hydraulic connectivity we used two methods:
(A) a linear regression, and (B) incorporating  hydraulic and geographic
connectivity into the meta-population model.  

In method (A) we fit a linear model to the median of the log of the
cross-neighborhood transmission coefficients ($\alpha_{j,i}$) from the fully
saturated model (model 3) using the hydraulic transition matrix and geographic
proximity matrix as covariates. The model can be written as follows:
$$y=\beta_0 + \beta_1x_1 + \beta_2x_2$$
where $y$ is a vector of the median of the log of the cross neighborhood
transmission coefficients ($\alpha_{i,j}$) and $x_1$ is a vector of hydraulic connectivity (Table S1) defined as 
$$x_1 \begin{cases} 0 & \text{ if no water connection exists } j \rightarrow i 
\\ 1 & \text{ if water connection exists} j \rightarrow i \end{cases}$$

and $x_2$ is a vector of geographic proximity (Table S2) defined as
$$x_2 \begin{cases} 0 &
\text{ if no shared border exists } j \rightarrow i \\ 1 & \text{ if shared
border exists } j \rightarrow i \end{cases}$$


In method (B) we expanded model 2 to allow the force of infection ($\alpha$) between two neighborhoods to vary depending
on if the neighborhoods are connected via water pipes such that
$$\alpha_{j,i}
\begin{cases} \alpha_{0} & \text{ if no water connection } j \rightarrow i \\
\alpha_{0} + \alpha_{1} & \text{ if water connection } j \rightarrow i 
\end{cases}$$
creating model 2b. We then expanded model 2b to incorporate geographic proximity
(model 2c) by adding an additional term $\alpha_{2}$ if the neighborhoods shared a
border. The resulting $\alpha_{i,j}$ can be described as
$$\alpha_{j,i} \begin{cases}
\alpha_{0} & \text{ if no shared border or water connection } j \rightarrow i \\
\alpha_{0} + \alpha_{1} & \text{ if no shared border but water connection }
j \rightarrow i \text{ exists}\\
\alpha_{0} + \alpha_{1} +\alpha_{2} & \text{ if shared border and water connection }
j \rightarrow i \text{ exists} 
\end{cases}$$ The effect of water, $\alpha_{1}$, and the effect of the shared border, 
$\alpha_{2}$, are not fitted to each neighborhood, but are shared citywide.

\newpage
## 1.3 Model fitting
The model used in the paper (model 3) was fit using `JAGS 3.4` and the `runjags` and
`rjags` libraries in R. The model priors were specified as thus:
$$^{new}\textrm{I}_{i,t+1}\sim Poisson \left ( \frac{S_{i,t}\phi}{N_{i}} (\beta_{i}I_{i,t} + \sum_{j\neq i}\alpha_{j,i}I_{j,t} ) \right )$$
$$log(\alpha_{j,i}) \sim N(\mu_{1}, \tau_{1})$$
$$log(\beta_{i}) \sim N(\mu_{2}, \tau_{2})$$
$$\mu_{1} \sim N(0, \frac{1}{0.001})$$
$$\mu_{2} \sim N(0, \frac{1}{0.001})$$
$$\tau_{1} \sim \Gamma(0.001, 0.001)$$
$$\tau_{2} \sim \Gamma(0.001, 0.001)$$
$$logit(\phi) \sim N(0, \frac{1}{0.001})$$ 
$$\gamma \sim exp(5)$$
The Gamma distribution for $\tau_1$ and $\tau_2$ was parameterized in terms of shape and rate. The exponential distribution for $\gamma$ was parameterized in terms of a rate.

## 1.4 Model selection  
We used the Watanabe-Akaike information criterion (WAIC) for model selection 
where a difference of at least 5 was considered significant. Note models 2.1 and
2.2 are not reported in the main text.


```{r WAIC}
source("functions/waic_dic_summary.R")
load(file = "Data/Rdata/waic_ls.Rdata")
mod_names <- c("model 1", "model 2", "model 2b", "model 2c",
               "model 2.1", "model 2.2", "model 3")


waic1 <- waic_list %>%
  waic_summary %>%
  data.frame() %>%
  `colnames<-`(mod_names)

pander(waic1)

```

For every realization of the epidemic, model 3 performed the best.

\newpage
```{r load summary}
load(file = "Data/Rdata/mcmc_summary_stats.Rdata")
```

## 1.5 Posterior summary statistics
For the selected model, model 3 (fully saturated model), we calculated the median and standard deviation of the posterior distribution for all fitted parameters.
The posterior median and standard deviation of $log(\beta_i)$ and $log(\alpha_{j,i})$ are in figure 1 and figure 2 respectively.  

The posterior median (standard deviation) for $\phi$ was `r round(mcmc_summary$phi_median, digits = 4)` (`r round(mcmc_summary$phi_sd, digits = 4)`).  
The posterior median (standard deviation) for $\gamma$ was `r round(mcmc_summary$gamma_median, digits = 4)` (`r round(mcmc_summary$gamma_sd, digits = 4)`)  



```{r posterior, fig.align="left"}
library(plotrix)

med_val_log <- round(mcmc_summary$betas_median_log, digits = 4)
sd_val_log <- round(mcmc_summary$betas_sd_log, digits = 4)

# Table plot of median posterior values
par(mar=c(5.5,6.5,1.5,2.9)) # Margins around plot ()
color2D.matplot(med_val_log, 
                show.values = 3,
                axes = FALSE,
                xlab = "",
                ylab = "",
                vcex = 0.7,
                vcol = "black",
                extremes = c("white", "white"))
xpos <- seq_len(ncol(med_val_log)) -0.3
ypos <- seq_len(ncol(med_val_log)) - 0.4
axis(3, # specifies top border x position
     at = xpos,
     labels = F, tick = FALSE, cex.axis = 0.7)
text(x = xpos,
     labels = names(med_val_log),
     srt = 45, # angle to rotate
     pos = 2, # specifies to put txt at top
     par("usr")[1] -.15, # 0.7 lines above the top. [4] places ref to top border
     adj = c(1,1),
     xpd = T) # not sure but allows txt to overflow table
axis(2, 
     at = ypos,
     labels = F, tick = FALSE, cex.axis = 0.7)
text(y = ypos,
     labels = rev(names(med_val_log)),
     srt = 45, # angle to rotate
     pos = 2, # specifies to put txt at top
     par("usr")[1] + -0.1,  # 0.1 lines left of left border. [1] places ref to left border
     adj = 0,
     xpd = T) # not sure but allows txt to overflow table
title(main = expression(paste("Posterior median of log(", beta[i], ") and log(", alpha['j,i'], ")")))
```

Figure 1: The posterior medians for $log(\beta_i)$ (diagonal) and $log(\alpha_{j,i})$ (off-diagonal) for each neighborhood from model 3. For example, row 4 (Kjoebmager), column 1 (Christianshavn) can be read as the posterior median of the log transmission coefficient for cases arising in Kjoebmager from cases in Christianshavn

```{r sd}
# Plot table of SD posterior values
par(mar=c(5.5,6.5,1.5,2.9)) # Margins around plot ()
color2D.matplot(sd_val_log, 
                show.values = 4,
                axes = FALSE,
                xlab = "",
                ylab = "",
                vcex = 0.7,
                vcol = "black",
                extremes = c("white", "white"))
xpos <- seq_len(ncol(sd_val_log)) -0.3
ypos <- seq_len(ncol(sd_val_log)) - 0.4
axis(3, # specifies top border x position
     at = xpos,
     labels = F, tick = FALSE, cex.axis = 0.7)
text(x = xpos,
     labels = names(sd_val_log),
     srt = 45, # angle to rotate
     pos = 2, # specifies to put txt at top
     par("usr")[1] -.15, # 0.7 lines above the top. [4] places ref to top border
     adj = c(1,1),
     xpd = T) # not sure but allows txt to overflow table
axis(2, 
     at = ypos,
     labels = F, tick = FALSE, cex.axis = 0.7)
text(y = ypos,
     labels = rev(names(sd_val_log)),
     srt = 45, # angle to rotate
     pos = 2, # specifies to put txt at top
     par("usr")[1] + -0.1,  # 0.1 lines left of left border. [1] places ref to left border
     adj = 0,
     xpd = T) # not sure but allows txt to overflow table
title(main = expression(paste("Posterior standard deviation of log(", beta[i], ") and log(", alpha['j,i'], ")")))

```

Figure 2: The posterior standard deviations for $log(\beta_i)$ (diagonal) and $log(\alpha_{j,i})$ (off-diagonal) for each neighborhood from model 3.
