---
title: "The importance of thinking beyond the water-supply in cholera epidemics: a historical urban case-study"
output: pdf_document
author: "Matthew Phelps, Andrew S. Azman, Joseph A. Lewnard, Marina Antillón, Lone Simonsen, Viggo Andreasen,  Peter K.M. Jensen, Virginia E. Pitzer."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')
pdf.options(encoding='ISOLatin2')
knitr::opts_knit$set(root.dir = normalizePath('..'))
knitr::opts_chunk$set(root.dir = normalizePath('..'))
library(tidyverse)
library(pander)
```

#Supplemental Text S1 
## 1.1 Model structure  
We constructed a series of nested models where
the force of infection, $\lambda$, acting upon a neighborhood was composed of an
external force of infection, $\alpha$, and an internal force of infection,
$\beta$.

From simple to complex we allowed (1) a single $\beta$ and single $\alpha$ for
all neighborhoods, (2) an individual $\beta_{i}$ for each neighborhood and a
single $\alpha$ for all neighborhoods, and (3) an individual $\beta_{i}$ for
each neighborhood and a single asymmetric $\alpha_{i,j}$ for each neighborhood
pair.


Two additional models were constructed where we relaxed the constraints on the
external forces of infection, $\alpha$. In model 2.1 allowed an individual
$\beta_{i}$ and $\alpha_{i}$ for each neighborhood and in model 2.2 we allowed
an individual $\beta_{i}$ for each neighborhood and symmetric $\alpha_{j,i}$ for
each neighborhood pair ($\alpha_{j,i} = \alpha_{i,j}$). These models were not
supported by the model selection process and were not reported in the main text.

Each model was based upon the following construction:
$$^{new}\textrm{I}_{i,t+1}\sim Poisson \left ( \frac{S_{i,t}\phi}{N_{i}} (\beta_{i}I_{i,t} + \sum_{j\neq i}\alpha_{j,i}I_{j,t} ) \right )$$


where:  
$^{new}\textrm{I}_{i,t}$ = the number of reported new infectious cases in each neighborhood $i$ at time $t$  
$I_{i,t}$ = the total number of infectious cases in each neighborhood $i$ at time $t$  
$S_{i,t}$ = the number of susceptible people in each neighborhood $i$ at time $t$  
$N_{i}$ = the total population of neighborhood $i$  
$\beta_{i}$ = the force of internal infection in neighborhood $i$  
$\alpha_{j,i}$ = the force of infection from neighborhood $j$ to neighborhood $i$.  


The total number of cases $I_{i,t}$ was updated via:
$$I_{i,t+1} = I_{i,t} + \frac{^{new}\textrm{I}_{i,t}}{\phi} - R_{i,t}$$  
where:  
$\phi$ = the fraction of cases that are reported  
$R_{i,t}$ = the number of people who recovered or died from infection.

The number of recovered individuals $R_{i,t}$ was updated via:  
$$R_{i,t} = \frac{1}{\gamma} I_{i,t}$$
where:  
$\gamma$ = the average duration of infectiousness (fixed at 5 days).  

The number of susceptible $S_{i,t}$ was updated via:
$$S_{i,t+1} = S_{i,t} - \frac{^{new}\textrm{I}_{i,t}}{\phi}$$  





## 1.2 Hydraulic connectivity and geographic proximity  
To assess the effect of hydraulic connectivity we used two methods:
(A) a linear regression, and (B) incorporating  hydraulic and geographic
connectivity into the meta-population model.  

In method (A) we fit a linear model to the median of the log of the
cross-neighborhood transmission coefficients ($\alpha_{i,j}$) from the fully
saturated model (model 3) using the hydraulic transition matrix and geographic
proximity matrix as covariates. The model can be written as follows:
$$y=\beta_0 + \beta_1x_1 + \beta_2x_2$$
where $y$ is a vector of the median of the log of the cross neighborhood
transmission coefficients ($\alpha_{i,j}$) and $x_1$ is a vector of hydraulic connectivity (Table S1) defined as 
$$x_1 \begin{cases} 0 & \text{ if no water connection exists } i \rightarrow j 
\\ 1 & \text{ if water connection exists} i \rightarrow j \end{cases}$$

and $x_2$ is a vector of geographic proximity (Table S2) defined as
$$x_2 \begin{cases} 0 &
\text{ if no shared border exists } i \rightarrow j \\ 1 & \text{ if shared
border exists } i \rightarrow j \end{cases}$$


In method (B) we expanded model 2 to allow the force of infection ($\lambda$) between two neighborhoods to vary depending
on if the neighborhoods are connected via water pipes such that
$$\lambda_{i,j}
\begin{cases} \alpha & \text{ if no water connection } i \rightarrow j \\
\alpha + \eta & \text{ if water connection } i \rightarrow j 
\end{cases}$$
creating model 2b. We then expanded model 2b to incorporate geographic proximity
(model 2c) by adding an additional term $\kappa$ if the neighborhoods shared a
border. The resulting $\lambda_{i,j}$ can be described as
$$\lambda_{i,j} \begin{cases}
\alpha & \text{ if no shared border or water connection } i \rightarrow j \\
\alpha + \eta & \text{ if no shared border but water connection }
i \rightarrow j \text{ exists}\\
\alpha + \eta +\kappa & \text{ if shared border and water connection }
i \rightarrow j \text{ exists} 
\end{cases}$$ The effect of water, $\eta$, and the effect of the shared border, 
$\kappa$, are not fitted to each neighborhood, but are shared citywide.


## 1.3 Model fitting
The model used in the paper (model 3) was fit using `JAGS 3.4` and the `runjags` and
`rjags` libraries in R. The model priors were specified as thus:
$$^{new}\textrm{I}_{i,t+1}\sim Poisson \left ( \frac{S_{i,t}\phi}{N_{i}} (\beta_{i}I_{i,t} + \sum_{j\neq i}\alpha_{j,i}I_{j,t} ) \right )$$
$$log(\alpha_{j,i}) \sim N(\mu_{1}, \tau_{1})$$
$$log(\beta_{i}) \sim N(\mu_{2}, \tau_{2})$$
$$\mu_{1} \sim N(0, \frac{1}{0.001})$$
$$\mu_{2} \sim N(0, \frac{1}{0.001})$$
$$\tau_{1} \sim \Gamma(0.001, 0.001)$$
$$\tau_{2} \sim \Gamma(0.001, 0.001)$$
$$logit(\phi) \sim N(0, \frac{1}{0.001})$$  

## 1.4 Model selection  
We used the Watanabe-Akaike information criterion (WAIC) for model selection 
where a difference of at least 5 was considered significant. Note models 2.1 and
2.2 are not reported in the main text.


```{r WAIC}
source("functions/waic_dic_summary.R")
load(file = "Data/Rdata/waic_ls.Rdata")
mod_names <- c("model 1", "model 2", "model 2b", "model 2c",
               "model 2.1", "model 2.2", "model 3")


waic1 <- waic_list %>%
  waic_summary %>%
  data.frame() %>%
  `colnames<-`(mod_names)

pander(waic1)

```

For every realization of the epidemic, model 3 performed the best.
